Refinements for Open Automata

### 2023-06-23 2021-Refinement/paper/refinementarticleV2.tex
meeting

There are not a lot of works that study refinement for open systems. Defining refinement of open systems as trace inclusion  is  addressed  as a notion of subtyping in type theory 
(e.g. \cite{GayH:2005,BravettiZ:2021}). The definition of refinement is based on a connection  between session types and communicating automata theories -- a notion of session automata
based on Communicating Finite-State Machines, that are used for  modelling processes communicating through FIFO channels.    The refinement of open systems is also defined in terms of  alternating simulation \cite{Alur:1998,deAlfaro:2021}.
Alternating simulation  is originating from the game theory \cite{deAlfaro:2003}, it allows  the study of relation between individual components by viewing them as alternating transition systems. In particular,  a refinement of game-based automata expresses that the refined component can offer more services (input actions) and fewer service demands (output actions). However, the composition of such automata may
lead to illegal states, where one automaton issues an output that is not acceptable as input in the other one. The theory of alternating simulation provides an optimistic approach to compute compatibility between automata based on the fact that each automaton expects the other to provide  legal inputs, i.e, two components can be composed if there is an environment where they can work together. Our approach  has some commonalities with  the above mentioned simulation \cite{deAlfaro:2021}: both are process-oriented approaches even if they are not based on the same notion of simulation, and both include in the model  how to compose and interact with processes that are accepted as parameters.
 \LUDO{\st{Je comprends pas l argument: ca a l air de dire qu on fait pas des systemes ouverts. en effet dans ce papier on decrit pas l environnement dans ce travail, a ce que je comprends? A REVOIR}}.

@@

The simulation  ensures the preservation of   
safety properties as deadlock-freeness  and, more generally, 
all linear temporal logic properties \cite{Abrialbooks96,Kouchnarenko:2007}.  Our focus is about the properties of simulation, in particular closure properties such as preservation under composition. Some works   have addressed the problem of non-introduction of blocking by composition. For example,  \cite{DIHEGO2020110598}  presented a refinement relation on process algebra based on failure semantics that is basically different from the (bi)simulation \cite{10.5555/640428.640430}. Indeed, failure semantics is based  on a notion of decorated traces.  Consequently, in this semantics,  refinement is defined in terms of inclusions between their decorated traces,  rather than by simulation. \LUDO{\st{the last sentence is not really informative}} \LUDO{\st{how do we compare our work?}}

@@

\begin{center}
\includegraphics[scale=0.7]{Figures/Thm1intuition}
\end{center}
\proof
Let us denote by $A_{12}$ (resp. $A_{13}$) the OA resulting from $A_1\subst{A_2}{k}$ (resp. $A_1\subst{A_3}{k}$),  to prove the theorem it is sufficient to prove that there exists a relation between states of the two OA that satisfies the conditions of the Definition \ref{Def:OA-Refinement}.

@@

Additionally, each transition $t_{13}$ of OT$(s_{13})$ is of the form $
g_1\land g_3 \land g$ with $g_1$ a guard of a transition $t_{1}$ of OT$(s_{1})$.
Thus have
\[
\neg{\bigvee_\subbox{t_1 \in \fOT{s_1}} \fguard{t_1}}  \implies \neg (\bigvee_\subbox{t_{13} \in \fOT{s_{13}}} \fguard{t_{13}}) \]

>>>>

The simulation  ensures the preservation of   
safety properties as deadlock-freeness  and, more generally, 
all linear temporal logic properties \cite{Abrialbooks96,Kouchnarenko:2007}.
The difference between the existing refinement principles have been studied in \cite{10.5555/640428.640430}, for example the authors explain in what sense failure semantics is different from  (bi)simulation in the compared systems and properties ensured. In this paper we particularly focus on the compositionality of simulation-based refinement.

There are not a lot of works that study refinement for open systems. Defining refinement of open systems as trace inclusion  is  addressed  as a notion of subtyping in type theory 
(e.g. \cite{GayH:2005,BravettiZ:2021}). The definition of refinement is based on a connection  between session types and communicating automata theories -- a notion of session automata
based on Communicating Finite-State Machines, that are used for  modelling processes communicating through FIFO channels.    The refinement of open systems is also defined in terms of  alternating simulation \cite{Alur:1998,deAlfaro:2021}.
Alternating simulation  is originating from the game theory \cite{deAlfaro:2003}, it allows  the study of relation between individual components by viewing them as alternating transition systems. In particular,  a refinement of game-based automata expresses that the refined component can offer more services (input actions) and fewer service demands (output actions). However, the composition of such automata may
lead to illegal states, where one automaton issues an output that is not acceptable as input in the other one. The theory of alternating simulation provides an optimistic approach to compute compatibility between automata based on the fact that each automaton expects the other to provide  legal inputs, i.e, two components can be composed if there is an environment where they can work together. Our approach  has some commonalities with  the above mentioned simulation \cite{deAlfaro:2021}: both are process-oriented approaches even if they are not based on the same notion of simulation, and both include in the model  how to compose and interact with processes that are accepted as parameters.
 \LUDO{TODO: what do we do better . in a different context?}.

@@

\LUDO{rewrite the following, relate non-blocking and the citation, is it the good one}
\LUDO{add reference to Olga for non blocking but she does not ensure compositionality in a very general sense as we do}
Some works   have addressed the problem of non-introduction of blocking by composition. For example,  \cite{DIHEGO2020110598}  presented a refinement relation on process algebra based on. Indeed, failure semantics is based  on a notion of decorated traces.  Consequently, in this semantics,  refinement is defined in terms of inclusions between their decorated traces,  rather than by simulation. \LUDO{\st{the last sentence is not really informative}} \LUDO{\st{how do we compare our work?}}

@@

\begin{center}
\includegraphics[scale=0.7]{Figures/Thm1intuition}
\end{center}
\proof
Let us denote by $A_{12}$ (resp. $A_{13}$) the OA resulting from $A_1\subst{A_2}{k}$ (resp. $A_1\subst{A_3}{k}$),  to prove the theorem it is sufficient to prove that there exists a relation between states of the two OA that satisfies the conditions of the Definition~\ref{Def:OA-Refinement}.

@@

Additionally, each transition $t_{13}$ of OT$(s_{13})$ is of the form $
g_1\land g_3 \land g$ with $g_1$ a guard of a transition $t_{1}$ of OT$(s_{1})$.
Thus we have
\[
\neg{\bigvee_\subbox{t_1 \in \fOT{s_1}} \fguard{t_1}}  \implies \neg (\bigvee_\subbox{t_{13} \in \fOT{s_{13}}} \fguard{t_{13}}) \]
